---
title: "Homework 1"
author: "lmuehlen"
date: "`r Sys.Date()`"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Setup

```{r message=FALSE, warning=FALSE}
library(tidyverse)
library(usethis)
# usethis::use_github()
set.seed(42)
```

# Functions

-   **generate_sample**

    -   generates 1 sample
```{r}
generate_sample <- function(N, y0 = 0, alpha = 0.8) {
  accumulate(rep(NA, N - 1), ~ alpha * .x + rnorm(1, 0, 1), .init = y0)
}
```

-   **draw_samples**

    -   draws [times] samples of sample sizes [N_values]
    
```{r}
draw_samples <- function(N_values, times, y0 = 0, alpha = 0.8) {
  map(N_values, ~ rerun(times, generate_sample(N = .x, y0, alpha)))
}
```

-   **estimate alpha**

    -   estimates the alpha

    -   easier for next steps to use dbl instead of list (list would be just map)
    
```{r}
estimate_alpha<- function(samples) {
  map_dbl(samples, ~ ar(.x,order.max = 1, aic = FALSE, method = "ols")$ar)
}
```


-   **compute_statistics**

    -   returns bias, variance and mse
    
```{r}
compute_statistics <- function(estimates, true_alpha = 0.8) {
  bias <- mean(estimates) - true_alpha
  variance <- var(estimates)
  mse <- variance + bias^2

  stats <- tibble(bias, variance, mse)

  return(stats)
}
```

-   **plot_statistics**

    -   plots bias, variance, and mse against sample size

```{r}
plot_statistics <- function(stats) {
  stats %>%
    ggplot(aes(x = as.numeric(N), y = value)) +
    geom_point() +
    geom_line(linetype = "dotted") +
    facet_wrap(~ factor(stat, levels = c("bias", "variance", "mse")), scales = "free_y", nrow = 3) +
    theme_minimal(base_size = 15) +
    theme(
      panel.spacing = unit(1, "lines"),
      strip.text = element_text(face = "bold")
    ) +
    labs(x = "Sample Size (N)", y = NULL)
}
```

# 3.1 Generate observations DGP

```{r}
true_alpha <- 0.8
sample_sizes <- c(10, 50, 100, 500, 1000)
samples <- draw_samples(sample_sizes, alpha = true_alpha, times = 5000)
alphas <- map(samples, estimate_alpha) %>% set_names(sample_sizes)
```

# 3.2 Compute and Plot Bias, Variance and MSE

## i)

```{r}
stats <- map(alphas, ~ compute_statistics(.x)) %>%
  bind_rows(.id = "N") %>%
  pivot_longer(-N, names_to = "stat", values_to = "value")
```

## ii)

```{r fig.height=7, fig.width=7}
plot_statistics(stats)
```

# 3.3 Interpretation

The MSE, variance, and bias converge in probability to 0 or in other words the estimator converges in probability to the true parameter value. Hence the OLS estimator is a consistent estimator for alpha.
