---
title: "ps2"
author: "lmuehlen"
date: "`r Sys.Date()`"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Packages

```{r message=FALSE}
library(readxl)
library(fixest)
library(texreg)
```

# Exercise 2
## 1

We use the Hausman test with\
$H_0$: educ is exogenous ($E[X^Tu]=E[W^Tu]=0$)\
$H_1$: educ is endogenous $E[X^Tu]\not=0$ $E[W^Tu]=0$

Manually
```{r}
wage2<-read_excel("Data/wage2.xlsx")

m_educ_sibs <- lm(educ ~ sibs, data=wage2)

wage2$res_educ_sibs <- m_educ_sibs$residuals

m_hausman <- lm(log(wage) ~ educ + exper + tenure + black + res_educ_sibs, data=wage2)

summary(m_hausman)

```

Quick
```{r}

iv_model<-feols(log(wage)~exper + tenure + black|educ~sibs,data = wage2)

fitstat(iv_model,~wh)

```

The p-value is 0.45. We cannot reject the hypothesis that educ is exogenous.

## 2i

**Relevance**: The instrument must be correlated with the endogenous explanatory variable, educ. For sibs, this seems plausible as the number of siblings might influence the allocation of financial resources within a family, potentially affecting the education level a child achieves. This aspect of relevance can be tested empirically by examining the strength and statistical significance of the instrument in the first stage of the 2SLS regression.

**Exogeneity**: The instrument must not be correlated with the error term in the regression equation. Empirically, exogeneity is challenging to test because it involves unobservable factors. We would argue that there could be potentially omitted variables affecting both the number of siblings and wages, such as parental education or socio-economic status. If such variables are not adequately controlled for, sibs might not satisfy the exogeneity criterion and, therefore, might not be a valid instrument.

## 2ii

```{r,results='asis'}
iv_model<-feols(log(wage)~exper + tenure + black|educ~sibs,data = wage2)
htmlreg(iv_model)

```

One additional year of education leads the an increase in monthly earnings by 9.36% on average ceteris paribus (if we assume that sib is a valid instrument).

## 3i

```{r, results="asis"}
m_stage1<-feols(educ ~ sibs+exper + tenure + black, data=wage2)

wage2$pre_educ<-predict(m_stage1)

m_stage2<-feols(log(wage) ~ pre_educ+exper + tenure + black, data=wage2)

htmlreg(list(m_stage1,m_stage2,iv_model),custom.coef.map=list("pre_educ"="educ","fit_educ"="educ","sibs"=NA,"Intercept"=NA,"exper"=NA,"tenure"=NA,"black"=NA),digits=4,custom.model.names = c("1st stage","2nd stage","Joint"))


```

## 3ii

The standard errors from the second stage regression when manually carrying out 2SLS are generally inappropriate since the estimation error of $\hat{X_i}$ is ignored.
